\LARGE \bf Preparation of Papers for IEEE Sponsored Conferences \& Symposia*

### 2017-06-12 appendix.tex
.

   ============
   The systems under test are assumed to be switched-mode to ease the
   presentation, but it is not a limitation of the presented
   technique. Black box systems which have hybrid dynamics due to
   multiple modes, are dealt in the same fashion.
   [TODO: $K$-refinement to differentiate the dynamics of different modes.
   We have added a benchmark to showcase this.
   [TODO: which one?]
   [TODO: clarify the absence of any controllers]
   [TODO: remove or clarify switched-mode systems]
   [TODO: add a hybrid benchmark]
   ============

   I imagine that the idea is to combine the generated discrete abstraction
   with a discrete controller model to verify both together. It would be
   good to state this clearly and to treat the resulting issues throughout
   the paper. I note that the experimental results do not treat this issue.
   Otherwise, why not change the title and focus the text on what you
   actually did?

   ============
   We will explore this in the future, but the current presentation
   only works with the plant, which is itself a hybrid dynamical
   system.
   [TODO: Add this in the conclusion/future work]
   ============

   (Similar remarks hold at a smaller scale. The discussion often describes
    what "can be" done, e.g., in sections 3.2 and 4.1. I would rather read
    about what "has been" done and what the result was!)

   ============
   [TODO]
   ============

@@

   My guess is that you later want to restrict the state-space by composing
   your abstraction with a controller model. Is that right? This probably
   seems obvious to you, but please put yourself in the place of someone who
   is new to the work and not an expert in the domain! In any case, this
   idea is not clearly stated, not addressed by the text, and not considered
   in the experimental evaluation.

   ============
   [TODO: I don't understand this comment, but it is related to
   point (iii) of the summary]
   Although we enumerate the abstract space, we require the model
   checker to find a concrete counter-example.
   ============

   Is this why the PWA transition system characterizes states by a valuation
   function v and not simply as tuples? If yes, how will this work? I do not
   understand why you have a valuation function when the system only has one
   variable (x).
   ============
   [TODO]
   ============
======================================
                            REVIEWER 2
======================================

@@

>> In general, finding the `optimal' number of PWA models can be
>> framed as a multiobjective optimization problem, where a small
>> number of clusters and an accurate affine map is desired. This is a
>> hard problem for arbitrary non-linear systems.
>> Hence, in this paper, we presented a scalable falsification approach
>> with a sub-optimal PWA identification approach.

@@

@@

@@

@@

>> As the choice of loss function and type of regression analysis
>> depends on the problem at hand, we show our methodology using the
>> simplest one. In practice, sophesticated statistical tests can be used
>> to find the a good analysis.

>>>>

>> The systems under test are assumed to be switched-mode to ease the
>> presentation, but it is not a limitation of the presented
>> technique. Black box systems which have hybrid dynamics due to
>> multiple modes, are dealt in the same fashion.

@@

>> BMC is required to find a concrete counter-example. The invariants
>> found through clustering (which has been since removed due to the
>> impending confusion) can not be used to conclude the presence of a
>> concrete coutner-examples by themselves as they are just an
>> approximation.

   My guess is that you later want to restrict the state-space by composing
   your abstraction with a controller model. Is that right?

>> This is a very important idea which we plan to explore in future
>> work. But it is not being presented in the paper as of now.

======================================
                            REVIEWER 2
======================================

@@

>> In general, finding the `optimal' number of PWA models can be
>> framed as a multi-objective optimization problem, where a small
>> number of clusters and an accurate affine map is desired. This is a
>> hard problem for arbitrary non-linear systems.
>> Hence, in this paper, we presented a scalable falsification approach
>> with a sub-optimal PWA identification approach.

@@

>> Current examples include generic dynamical systems. But the same
>> procedure can be used to find falsifications in closed loop control
>> systems. We plan to do extensive case studies in the future.

@@

>> In earlier works, S3CAM was shown to be better than S-Taliro.
>> Hence, for now only S3CAM was compared. In the future we plan to have
>> extensive case-studies and comparison with other tools.

@@

>> We tried polynomial kernels but Z3 can not scale even on small
>> number of polynomial constraints. Instead we used non-linear
>> optimization solvers. But, they were very sensitive to the initial
>> guess.

@@

>> As the choice of loss function and type of regression analysis
>> depends on the problem at hand, we show our methodology using the
>> simplest one. In practice, sophisticated statistical tests can be used
>> to find the a good analysis.

### 2017-06-12 implementation.tex
.

PW-Rel is implemented in Python. It uses S3CAM to generate the graph
abstraction and regression routines from
Scikit-learn~\cite{pedregosa2011scikit} to compute
$\RTSAbstraction{\epsilon}{\System}$ of the given system $\System$.
It then constructs relational abstraction and uses a bounded model
checker to find a counter-example to the given property in
$\RTSAbstraction{\epsilon}{\System}$. Finally, using simulations, it
confirms if the counter-example is valid for $\System$, in which case
a successfull falsification run is concluded.

To check the transition system, PW-Rel can either invoke symbolic
bounded model checkers or explicit model checkers. For symbolic BMC,
the off-the-shelf SAL modelchecker~\cite{SAL-SRI} is used. It is
configured to encode the problem as a single query for a given depth
$k$, where $k$ is the bound on the length of the counter-example. It
was computed by the length of the abstract counter-examples found. SAL
uses the SMT solver Yices2~\cite{dutertre2014yices} to solve the
encoded SMT constraints.

@@

Explicit model checking was implemented out using a naive DFS
(depth-first-exploration) stratergy, wherein, each abstract path is
enumerated and checked for feasibility using a an optimization
routine. Intefaces to various solvers were implemented, including GLPK
and Ipopt. If feasible, the returned concrete counter-example is
simulated and checked for existance in $\System$. Also, PW-Rel
extracts the bounds on the initial parameters by formulating multiple
optimization problems. The bounds can be regarded as an over-appr

% If the existance check fails, multiple
% parameter values can be sampled within the determined bounds to search
% for counter-examples in the neighbhourhood of the returned concrete
% counter-example.

@@

We benchmarked our implementation and compared it with S3CAM.  The
preliminary results are tabulated in~\tabref{res-rel}. We used few of
the benchmarks described in \cite{zutshi2014multiple}, including the
Van der Pol oscillator, Brusselator, Lorenz attractor and the
Navigation benchmark. We also tried PW-Rel on the insulin
benchmark~\cite{chenformal} and lane change maneuver
benchmark~\cite{althoff2012reachability}. PW-Rel was compared with
S3CAM by running both $10$ times with different seeds (as both use
randomized algorithms), and their running times were averaged. PW-Rel
was evluated with both the symbolic BMC SAL (PW-Rel+SAL) and the
explicit BMC (PW-Rel+OPT). %\todo{SatEx was also tried, but ...}.
We
tabulate both the total time taken by PW-Rel+SAL and PW-Rel+OPT to
compute the counter-example and compare against an implementation of
S3CAM.  Different abstraction parameters are used for S3CAM and
PW-Rel, due to the differences in which they operate. PW-Rel
requires a finer abstraction to begin with in order to minimize the
modeling errors. On the other hand, a coarser abstraction as the
initial abstraction works better for S3CAM due to the expensive
refinement procedure involved.~\footnote{In all our experiments we
used $1$-relational modeling to create the transition system.
This was determined to be a good trade-off between longer simulations
and increase in precision, with respect to the mentioned benchmarks.}

\begin{table}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in minutes is noted under
    \textbf{PW-Rel+BMC solver} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
    Benchmark & PW-Rel + SAL & PW-Rel + LP & S3CAM\\
\midrule
    Van der Pol ($\prop_3$)   &$0.2$ & $0.34$ & $0.4$\\
    Brusselator               &$0.06$ & $0.06$ & $0.03$\\
    Lorenz                    &$0.4$ & $0.33$ & $0.58$\\
%B.Ball                    &  &   & \\
    Nav ($\prop_P$)           &$TO$ & $28.5$  & $10$ \\%9423 + 4480
    Nav ($\prop_Q$)           &$TO$ & $26$   & $9.1$ \\%8105 + 970
    Nav ($\prop_R$)           &$TO$ & $39.2$  & $33.3$\\
    Nav ($\prop_S$)           &$TO$ & $71$  & $35$\\
    Bouncing Ball             &$Safe$ & $Safe$  & $7.5$\\

    insulin-st-1             &$Safe$ & $Safe$  & $0.0$\\
    insulin-st-2             &$Safe$ & $Safe$  & $0.0$\\
    lane change              &$Safe$ & $Safe$  & $0.0$\\
    HVAC                     &$Safe$ & $Safe$  & $0.0$\\

@@

However, we need to explore more benchmarks to be conclusive on the
tool's performance on real world examples. Although, PW-Rel does not
need to refine like S3CAM, and should be more scalable, it is still
constrained by the accuracy of the piecewise models. Its performance can be
explained by the difficulty in finding an abstraction
($\quant_\epsilon$), which is fine enough, to result in accurate
prediction models but coarse enough to be manageable by the current
bmc tools.  To obtain a reproducible counter-example from the model
checker (and avoid false positives), the transition system should be
created from models with high accuracy, for which a finer abstraction
is required.  However, the exploration of a finer abstraction is
exponentially more complex and results in a much bigger transition
system, which the current state of the art bounded model checkers
(which use generic SMT solvers) can not handle under reasonable
resources.

>>>>

PW-Rel is implemented in Python. It uses S3CAM to generate the graph
abstraction and regression routines from
Scikit-learn~\cite{pedregosa2011scikit} to compute
$\RTSAbstraction{\epsilon}{\System}$ of the given system $\System$.
It then constructs relational abstraction and uses a bounded model
checker to find a counter-example to the given property in
$\RTSAbstraction{\epsilon}{\System}$. Finally, using simulations, it
confirms if the counter-example is valid for $\System$, in which case
a successful falsification run is concluded.

To check the transition system, PW-Rel can either invoke symbolic
bounded model checkers or explicit model checkers. For symbolic BMC,
the off-the-shelf SAL model checker~\cite{SAL-SRI} is used. It is
configured to encode the problem as a single query for a given depth
$k$, where $k$ is the bound on the length of the counter-example. It
was computed by the length of the abstract counter-examples found. SAL
uses the SMT solver Yices2~\cite{dutertre2014yices} to solve the
encoded SMT constraints.

@@

Explicit model checking was implemented out using a naive DFS
(depth-first-exploration) strategy, wherein, each abstract path is
enumerated and checked for feasibility using a an optimization
routine. Interfaces to various solvers were implemented, including GLPK
and Ipopt. If feasible, the returned concrete counter-example is
simulated and checked for existence in $\System$. Also, PW-Rel
extracts the bounds on the initial parameters by formulating multiple
optimization problems. The bounds can be regarded as an over-appr

% If the existence check fails, multiple
% parameter values can be sampled within the determined bounds to search
% for counter-examples in the neighborhood of the returned concrete
% counter-example.

@@

We benchmarked our implementation and compared it with S3CAM.  The
preliminary results are tabulated in~\tabref{res-rel}. We used few of
the benchmarks described in \cite{zutshi2014multiple}, including the
Van der Pol oscillator, Brusselator, Lorenz attractor and the
Navigation benchmark.
%We also tried PW-Rel on the insulin
%benchmark~\cite{chenformal} and lane change maneuver
%benchmark~\cite{althoff2012reachability}.
PW-Rel was compared with
S3CAM by running both $10$ times with different seeds (as both use
randomized algorithms), and their running times were averaged. PW-Rel
was evaluated with both the symbolic BMC SAL (PW-Rel+SAL) and the
explicit BMC (PW-Rel+OPT). %\todo{SatEx was also tried, but ...}.
An affine kernel was used for modeling and the linear programming
solver GLPK was used for checking feasibility of the constraints.
We also tried polynomial kernels, and used Z3 SMT
solver~\cite{DeMoura+Bjorner/08/Z3} to check the symbolic BMC.
Unfortunately, Z3 does not scale well on polynomial constraints and
consistently timed out. For explicit BMC, several non-linear
optimization engines were tried, but the results were inconsistent due
to sensitivity towards the initial guess.

We tabulate both the total time taken by PW-Rel+SAL and PW-Rel+OPT to
compute the counter-example and compare against an implementation of
S3CAM.  Different abstraction parameters are used for S3CAM and
PW-Rel, due to the differences in which they operate. PW-Rel
requires a finer abstraction to begin with in order to minimize the
modeling errors. On the other hand, a coarser abstraction as the
initial abstraction works better for S3CAM due to the expensive
refinement procedure involved.~\footnote{In all our experiments we
used $1$-relational modeling to create the transition system.
This was determined to be a good trade-off between longer simulations
and increase in precision, with respect to the mentioned benchmarks.}

\begin{table}[!htbp]
\centering
\caption{Avg. timings for benchmarks. The \textbf{BMC} column lists
time taken by the BMC engine. The total time in minutes is noted under
    \textbf{PW-Rel+BMC solver} and \textbf{S3CAM}.
$TO$ signifies time $>5hr$, after which the search was killed.}
\label{tab:res-rel}
\begin{tabular}{@{}llll@{}}
\toprule
    Benchmark & PW-Rel + SAL & PW-Rel + OPT & S3CAM\\
\midrule
    Van der Pol ($\prop_3$)   &$0.2$ & $0.34$ & $0.4$\\
    Brusselator               &$0.06$ & $0.06$ & $0.03$\\
    Lorenz                    &$0.4$ & $0.33$ & $0.58$\\
%B.Ball                    &  &   & \\
    Nav ($\prop_P$)           &$TO$ & $28.5$  & $10$ \\%9423 + 4480
    Nav ($\prop_Q$)           &$TO$ & $26$   & $9.1$ \\%8105 + 970
    Nav ($\prop_R$)           &$TO$ & $39.2$  & $33.3$\\
    Nav ($\prop_S$)           &$TO$ & $71$  & $35$\\
    Bouncing Ball             &$Safe$ & $Safe$  & $7.5$\\

    %insulin-st-1             &$Safe$ & $Safe$  & $0.0$\\
    %insulin-st-2             &$Safe$ & $Safe$  & $0.0$\\
    %lane change              &$Safe$ & $Safe$  & $0.0$\\
    %HVAC                     &$Safe$ & $Safe$  & $0.0$\\

@@

However, we need to explore more benchmarks to be conclusive on the
tool's performance on real world examples. Although, PW-Rel does not
need to refine like S3CAM, and should be more scalable, it is still
constrained by the accuracy of the piecewise models. Its performance can be
explained by the difficulty in finding an abstraction
($\quant_\epsilon$), which is fine enough, to result in accurate
prediction models but coarse enough to be manageable by the current
BMC tools.  To obtain a reproducible counter-example from the model
checker (and avoid false positives), the transition system should be
created from models with high accuracy, for which a finer abstraction
is required.  However, the exploration of a finer abstraction is
exponentially more complex and results in a much bigger transition
system, which the current state of the art bounded model checkers
(which use generic SMT solvers) can not handle under reasonable
resources.

### 2017-06-12 learning.tex
.

In statistics, parametric regression is the problem of finding the
paremters of a function or \textit{predictor}, which can `best'
summarize the relationship between the given set of observed input
$\x$ and output $\x'$ vectors. The basis functions used for the
predictor are fixed beforehand and the paramters are searched for.
The notion of `best' is formally captured using a loss function which
can be non-linear. We use the commonly used loss function
\textit{ordinary least squares} (OLS) to fit affine and polynomial
kernels of fixed degree.  ~\footnote{The choice of loss function and
regression analysis affects the quality of the learnt models. As the
procss of selecting an appropriate heuristic is not straightforward,
we just use the simplest one.} OLS defines its loss function as the
sum of squares of the errors in prediction. The task is then to
determine the parameters, such that the least square error of the
predictor is minimized for the given data set.  The details on
performing linear regressoin can be found in standard texts on
learning and statistics~\cite{friedman2001elements}.

@@

>>>>

In statistics, parametric regression is the problem of finding the
parameters of a function or \textit{predictor}, which can `best'
summarize the relationship between the given set of observed input
$\x$ and output $\x'$ vectors. The basis functions used for the
predictor are fixed beforehand and the parameters are searched for.
The notion of `best' is formally captured using a loss function which
can be non-linear. We use the commonly used loss function
\textit{ordinary least squares} (OLS) to fit affine and polynomial
kernels of fixed degree.  ~\footnote{The choice of loss function and
regression analysis affects the quality of the learnt models. As the
process of selecting an appropriate heuristic is not straightforward,
we just use the simplest one.} OLS defines its loss function as the
sum of squares of the errors in prediction. The task is then to
determine the parameters, such that the least square error of the
predictor is minimized for the given data set.  The details on
performing linear regression can be found in standard texts on
learning and statistics~\cite{friedman2001elements}.

@@

\begin{algorithm}[t]
\DontPrintSemicolon
\caption{RelationalAbstraction\label{algo:relabs}}
\KwIn{$\simmap_\Delta$, $\TSAbstraction{\epsilon}{\System}$,
    regression kernel $F_k$}
\KwOut{$\RTSAbstraction{\epsilon}{\System}$}
    \ForAll{$(\Cells,\Cells') \in \CellEdges$}{
    $F, \delta = \mathsf{Regression(F_k, \simmap_\Delta, (\Cells,\Cells'))}$ \;
    $r(\vx,\vx^+) = F, \delta$
    %$\RTSAbstraction{\epsilon}{\System}$ $\assign$
    %$\mathsf{k-RelationalAbstraction}$($\TSAbstraction{\epsilon}{\System}, F$) \nllabel{algoline:relabs} \;
    %$\Pi$ $\assign$ $\mathsf{BoundedModelCheck}$($\RTSAbstraction{\epsilon}{\System}$) \nllabel{algoline:bmc} \;
    %\lIf {$\Pi = \emptyset$}{ return FAIL \nllabel{algoline:nocex} }
    %\lIf {$\mathsf{checkIfConcretizable}(\Pi)$}{ return VIOLATION
    %\nllabel{algoline:checkspurious} }
}
\end{algorithm}

### 2017-06-12 prelims2.tex
.

In statistics, linear regression is the problem of finding an affine
function or \textit{predictor}, which can `best' summarize the
relationship between the given set of observed input $\x$ and output
$\x'$ vectors. The notion of `best' is formally captured using a loss
function (which can be non-linear), resulting in different kinds of
regression analysis including ridge, lasso, huber, generelized and
many other regressions. For simplicity, we present the commonly used
loss function \textit{ordinary least squares} (OLS) in the next
section.
~\footnote{The choice of regression analysis should ideally be based on the
quality of data. As we assume the data (from the simulator) to be
noise free, the affecte on result of the falsification search is
minimal.
Though it is important,  not central to the
discussion as long as the learnt models  can be used to learn different
models}.

>>>>

In statistics, linear regression is the problem of finding an affine
function or \textit{predictor}, which can `best' summarize the
relationship between the given set of observed input $\x$ and output
$\x'$ vectors. The notion of `best' is formally captured using a loss
function (which can be non-linear), resulting in different kinds of
regression analysis including ridge, lasso, huber, generalized and
many other regressions. For simplicity, we present the commonly used
loss function \textit{ordinary least squares} (OLS) in the next
section.
~\footnote{The choice of regression analysis should ideally be based on the
quality of data. As we assume the data (from the simulator) to be
noise free, the affecte on result of the falsification search is
minimal.
Though it is important,  not central to the
discussion as long as the learnt models  can be used to learn different
models}.

### 2017-06-12 scam.tex
.

We assume that the system-under-test is a hybrid dynamical system, \ie, a
system which has discrete modes and a continuous state-space, and which
evolvution is either (1) continuous in time according to the dynamics
associated with its current discrete mode (typically described using
differential equations over its continuous state), or (2) through
discrete jumps that possibly change the discrete mode of the system and
possibly reset its continuous state to some value.  However, we assume
that the system is {\em effectively black-box}. This means that we do not
assume any knowledge of the symbolic dynamical equations describing the
system behavior. We do assume some knowledge of the system: an {\em
interface} view, in which we are allowed to stimulate the system with
inputs and observe its outputs.

We formally define a system-under-test $\System$ as a tuple $(\Modes,
\ContStates, \InitModes, \InitStates, \Delta, \Inputs, \Outputs)$.
Here, system state-space (denoted $\HybridStates$) is a subset of
$\Modes \times \ContStates$, where $\Modes$ is a finite set of
discrete modes, and $\ContStates$ is a compact subset of $\Reals^n$
representing the set of continuous states.  The sets $\InitModes
\subseteq \Modes$ and $\InitStates \subseteq \ContStates$ represent
initial modes and states respectively.  We assume that the system is
{\em initialized} to some initial mode $\initmode \in \InitModes$ and
an initial state $\initstate \in \InitStates$ at time $t=0$.  We
assume that the system has $m$ exogenuous inputs which take values
from the set $\Inputs$ which is a compact subset of $\Reals^m$. We
assume that the system has $\ell \le n$ outputs, where $\Outputs
\subseteq \HybridStates$. In this paper, we assume that the entire
system state is observable.  In other words, $\Outputs =
\HybridStates$.

@@

We remark that our abstraction is {\em neither an overapproximation
nor an underapproximation} of the continuous-time continuous-space
reachability relation between states. The purpose of the $\epsilon$-TS
is to discover falsifying trajectories. This can be accomplished by
Algorithm~\ref{algo:s3camBMC} that uses CounterExample-Guided
Abstraction and Refinement (CEGAR).

>>>>

We assume that the system-under-test is a hybrid dynamical system, \ie, a
system which has discrete modes and a continuous state-space, and which
evolution is either (1) continuous in time according to the dynamics
associated with its current discrete mode (typically described using
differential equations over its continuous state), or (2) through
discrete jumps that possibly change the discrete mode of the system and
possibly reset its continuous state to some value.  However, we assume
that the system is {\em effectively black-box}. This means that we do not
assume any knowledge of the symbolic dynamical equations describing the
system behavior. We do assume some knowledge of the system: an {\em
interface} view, in which we are allowed to stimulate the system with
inputs and observe its outputs.

We formally define a system-under-test $\System$ as a tuple $(\Modes,
\ContStates, \InitModes, \InitStates, \Delta, \Inputs, \Outputs)$.
Here, system state-space (denoted $\HybridStates$) is a subset of
$\Modes \times \ContStates$, where $\Modes$ is a finite set of
discrete modes, and $\ContStates$ is a compact subset of $\Reals^n$
representing the set of continuous states.  The sets $\InitModes
\subseteq \Modes$ and $\InitStates \subseteq \ContStates$ represent
initial modes and states respectively.  We assume that the system is
{\em initialized} to some initial mode $\initmode \in \InitModes$ and
an initial state $\initstate \in \InitStates$ at time $t=0$.  We
assume that the system has $m$ exogenous inputs which take values
from the set $\Inputs$ which is a compact subset of $\Reals^m$. We
assume that the system has $\ell \le n$ outputs, where $\Outputs
\subseteq \HybridStates$. In this paper, we assume that the entire
system state is observable.  In other words, $\Outputs =
\HybridStates$.

@@

We remark that our abstraction is {\em neither an over-approximation
nor an under-approximation} of the continuous-time continuous-space
reachability relation between states. The purpose of the $\epsilon$-TS
is to discover falsifying trajectories. This can be accomplished by
Algorithm~\ref{algo:s3camBMC} that uses CounterExample-Guided
Abstraction and Refinement (CEGAR).
